---
title: "Automating Reports with phstemplates"
author: 'Jack Hannah'
date: "11 May 2023 <br><br> Slides: https://github.com/jackhannah95/presentations/tree/main/2023-05-11_phstemplates"
output:
  ioslides_presentation:
    css: my-theme.css
---

```{r setup, include = FALSE}
# Load packages required
library(knitr)
library(here)

opts_chunk$set(echo = TRUE)
```

## Traditional Publication Process

```{r out.width = "100%", fig.align = 'center', echo = FALSE}
include_graphics(here("2023-05-11_phstemplates", 
                      "images", 
                      "traditional_process.png"))
```

## R Markdown

* Document format within RStudio.

* Allows code chunks to be embedded within text.

* Supports a variety of output formats, including Word, PDF, PowerPoint and HTML.

## R Markdown In Practice

This is from the main points in the [2023-04-04 Cancer Waiting Times summary report](https://publichealthscotland.scot/media/18856/2023-04-04-cwt-summary.pdf):

_4,262 eligible patients were referred for the 62-day standard, an increase of 2.3% from the previous quarter, and an increase of 14.5% from the quarter ending 31 December 2019._

## R Markdown In Practice

Three numeric values required to create this sentence:

```{r}
# Eligible patients in current quarter
n_current <- 4262

# Eligible patients in previous quarter
n_previous <- 4161

# Eligible patients in last pre-covid quarter
n_pre_covid <- 3722
```

## R Markdown In Practice

`prettyNum(n_current, big.mark = ",")` eligible patients were referred for the 62-day standard, `ifelse(n_current > n_previous, "an increase", "a decrease")` of `scales::percent((n_current - n_previous) / n_previous, accuracy = 0.1)` from the previous quarter, and `ifelse(n_current > n_pre_covid, "an increase", "a decrease")` of `scales::percent((n_current - n_pre_covid) / n_pre_covid, accuracy = 0.1)` from the quarter ending 31 December 2019.

**`r prettyNum(n_current, big.mark = ",")`** eligible patients were referred for the 62-day standard, **`r ifelse(n_current > n_previous, "an increase", "a decrease")`** of **`r scales::percent((n_current - n_previous) / n_previous, accuracy = 0.1)`** from the previous quarter, and **`r ifelse(n_current > n_pre_covid, "an increase", "a decrease")`** of **`r scales::percent((n_current - n_pre_covid) / n_pre_covid, accuracy = 0.1)`** from the quarter ending 31 December 2019.

## phstemplates

* A package designed to provide standard R templates for PHS.

* Includes templates for code structure (scripts, projects, shiny apps) and output formats (Official/National Statistics reports, PowerPoint, HTML slides).

* Can be installed on the Workbench with `install.packages("phstemplates")`.

## Using phstemplates

To create (e.g.) a summary report for a National Statistics publication:

* File -> New File -> R Markdown... -> From Template -> PHS National Stats Report

```{r out.width = "40%", fig.align = 'center', echo = FALSE}
include_graphics(here("2023-05-11_phstemplates", 
                      "images", 
                      "ns_template.png"))
```

## Using phstemplates

* Create your report inside the `.Rmd` file.

* `Knit` the report:

```{r out.width = "100%", fig.align = 'center', echo = FALSE}
include_graphics(here("2023-05-11_phstemplates", 
                      "images", 
                      "knit.png"))
```

* [View the end result](https://github.com/jackhannah95/presentations/raw/main/2023-05-11_phstemplates/example-report/example-report.docx).

## Links and Resources

* [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)

* [phstemplates GitHub repository](https://github.com/Public-Health-Scotland/phstemplates)

* [End of Life publication GitHub repository](https://github.com/Public-Health-Scotland/end-of-life-pub/)
